---
title: "Assignment-2"
author: "Neha Bais"
date: "9/22/2020"
output: html_notebook
---

# Use the PubMed dataset and download at least five years of scientific articles.


Required libraries
```{r}
#install.packages("RISmed")
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("ggthemes")
#install.packages("extrafont")
#install.packages("plyr")
#install.packages("ggalt")
#install.packages('tidytext')
#install.packages("mapproj")
#install.packages("ggmap")


library(RISmed)
library(tidyverse)
library(dplyr)
library(tidytext)
library(ggplot2)
library(ggthemes)
library(extrafont)
library(plyr)
library(ggalt)
library(ggmap)
library(maps)
library(mapproj)
```

# Part-1
####### RISmed library is used for this task.
####### I restricted 1000 articles for each year 

```{r}
year_list <- c(2016,2017,2018,2019,2020)
datalist <- list()

for (year in year_list){
  search_topic <- ''
  search_query <- EUtilsSummary(search_topic, retmax=1000, mindate=year, maxdate=year)
  summary(search_query)
  
  # see the ids of our returned query
  QueryId(search_query)
  
  # get actual data from PubMed
  records<- EUtilsGet(search_query)
  class(records)
  
  # Extract required fields only
  abstracts <- data.frame(title = records@ArticleTitle,
                          year = records@YearPubmed)
  
  # add data to list
  datalist[[year]] <- abstracts
  
}

# Combine dataframe
big_data <- do.call(rbind, datalist)
head(big_data, 3)
```

##### Calculating frequencies for each of the following terms: Influenza, Obesity, Cancer, Covid-19 

```{r}
#######################
#  Getting word freq  #
#######################

words_count <- big_data %>% 
               unnest_tokens(input=title, output=word,
                            format = 'text', token = 'words', drop=FALSE, 
                            to_lower=TRUE,strip_punct=TRUE) %>%
               dplyr::count(year, word)

desired_words <- words_count[grep('influenza|cancer|obesity|covid',words_count$word),]

# Getting results of my words
final_result <- list()
my_words <- c('influenza','cancer','obesity','covid')

for(word in my_words){
  df <- desired_words[desired_words$word==word,]
  final_result[[word]] <- df
}

final_result <- do.call(rbind, final_result)
rownames(final_result) <- NULL
colnames(final_result) <- c('year', 'word', 'freq')
print(final_result, row.names = F)
```

### a. Area Plot comparing frequencies for the defined terms for each year.

```{r}
###############
#  Area Plot  #
###############
p1 <- ggplot() + 
  geom_area(aes(y = freq, x = year, fill = word), data = final_result, stat="identity")+
  ggtitle("Area Plot for Word Frequency") +
  theme(plot.title = element_text(hjust = 0.5)) 
    
p1

```

### b. Dumbell chart 
##### Finding frequencies of below keywords for 2019 and 2020 years:
##### Influenza, Covid-19, Depression, Mental health, Physical activity, Wearable

```{r}
# selecting records of only 2019 and 2020
reduced_df <- words_count[words_count$year %in% c(2019,2020), ]
table(reduced_df$year)

# search for desired words in all words count dataframe
desired_words2 <- reduced_df[grep('influenza|depression|mental health|covid|physical activity|wearable',
                                  reduced_df$word),]

# Getting results of my words
words_dumb_chart <- list()
words_dc <- c('influenza','depression','mental health','covid','physical activity','wearable')

for(word in words_dc){
  df2 <- desired_words2[desired_words2$word==word,]
  words_dumb_chart[[word]] <- df2
}

words_dumb_chart <- do.call(rbind, words_dumb_chart)
rownames(words_dumb_chart) <- NULL
print(words_dumb_chart, row.names = F)


w <- c('influenza','depression','mental health','covid','physical activity','wearable')
c_2019 <- c(0,3,0,0,0,0)
c_2020 <- c(4,12,0,49,0,0)
df_dchart <- data.frame(word = w, count_2019=c_2019, count_2020=c_2020)
```

###  Dumbell Plot comparing changes of defined keywords for 2 years.
```{r}

p2 <- ggplot(df_dchart, aes(x=count_2019, xend=count_2020, y=word)) + 
  geom_dumbbell() +
  ggtitle("Dumbell plot for word frequency 2019 vs 2020") +
  theme(plot.title = element_text(hjust = 0.5)) 

p2
```
The above plot shows that only depression term appears in both 2019 and 2020 years and the change in frequency is from 3 to 12.


# Part-2
#### Create a choropleth map for visualising number of COVID comfirmed cases in Massachusetts.
#### https://www.mass.gov/info-details/covid-19-response-reporting

```{r}
# State data
s <- map_data('county')
state_data <- s %>% filter(region == 'massachusetts')

# COVID data - USA
covid_data <- read.csv('mass_covid_cases.csv', header = T)


# Merge state geographical data and covid data
covid_data$County <- tolower(covid_data$County)
final_covid_data  <- merge(state_data, covid_data,
                     by.x = 'subregion',
                     by.y = 'County')

#data <- data[order(data$Confirmed.cases.till.date), ]

#aggregate data to get mean latitude and mean longitude for each county
cnames <- aggregate(cbind(long, lat) ~ subregion, data=final_covid_data, FUN=function(x) mean(range(x)))

head(final_covid_data)
```


### Plot the covid cases data on Massachusetts map

###### Barnstable    	1697
###### Berkshire	      702
###### Bristol       	9700
###### Dukes         	61
###### Essex	          18872
###### Franklin	      401
###### Hampden	        7947
###### Hampshire     	1203
###### Middlesex	      26545
###### Nantucket     	85
###### Norfolk	        9980
###### Plymouth      	9487
###### Suffolk	        23552
###### Worcester     	14007


```{r}
# Plot the data on massachusetts map
ggplot() +
  geom_polygon(data = final_covid_data, 
               aes(x = long, y = lat, group = group, fill = Confirmed.cases.till.date), 
              color = 'black', size = 0.5) +
  coord_map()+
  scale_fill_distiller(name="Confirmed cases", palette  = 'Reds', direction = 1)+
  theme_nothing(legend = TRUE) +
  ggtitle("Covid-19 confirmed cases in Massachusetts") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(data=cnames, aes(long, lat, label = subregion), size=3, fontface="bold")
```

